Using Hybrid Machine Learning for Fundamental Stellar Parameter Inference

### 2024-04-09 paper/paper2_neo/paper.tex
Auto stash before merge of "working" and "origin/working"

\section{Introduction}
Text

@@

\begin{table}[h]
    \hspace*{-1.1cm}\begin{tabular}{|c|c|c|}
        \hline
        \multicolumn{3}{|c|}{\textbf{PHOENIX Grid Subset}}\\
        \hline\hline
        Parameter & Symbol & Range\\
        \hline
        Alpha Element Abundance & $\alpha$ & [0] dex\\
        Iron Abundance & [Fe/H] & [-0.5, 0.5] dex\\
        Effective Temperature & $T_{\mathrm{eff}}$ & [2300, 12000] K\\
        Surface Gravity & $\log(g)$ & [0, 5]\\
        Wavelength & $\lambda$ & [8050, 12850] \AA\\
        \hline
    \end{tabular}
    \caption{The subset of the PHOENIX grid used in this study. These limits 
    were imposed to reduce the computational cost of the algorithms. The 
    wavelength limits in particular roughly line up with that of the 
    Habitable Zone Planet Finder (HPF) spectrograph.}
\end{table}

@@

Once optimization was complete, the list of identified lines, present in the
state dictionary of \texttt{blase}'s model, was saved to a \texttt{.pt} file.

@@

\subsection{Continuously Evaluable Manifolds}
For each line, the inputs to the interpolator were the three input parameters $T_{\mathrm{eff}}$,
$\log(g)$, and $[Fe/H]$, and the output was a list of four parameters, 
$\delta\mu$, $\ln(a)$, $\sigma$, and $\gamma$. For each line, one of these 
interpolator objects was created using linear interpolation, and these 
interpolators were aggregated into a single list, which was when pickled 
and written to a \texttt{.pkl} file. These interpolators represent multiple 
manifolds mapping fundamental stellar parameters to spectral line properties. 
These interpolators could now be evaluated at any point lying within the 
domain of the PHOENIX subset, turning a discretely sampled PHOENIX subset 
into a continuous one. With the given size of the PHOENIX subset, the 
interpolator list takes up 19 GB of disk space.

\section{Bayesian Optimization}
Text

\section{Inference Testing}
\subsection{Test Data}
Text

@@

\section{Discussion}

@@

>>>>

\section{Introduction}
Stellar spectra are exceedingly rich sources of information about the stars
which create them. They contain information about the fundamental properties
of the star such as its temperature and surface gravity, as well as
countless absorption lines reflecting the star's chemical composition. On 
top of that, extrinsic factors such as the radial velocity of the star or 
the rotation of the star itself further transform spectra. When these spectra
are observed, even the location of the observing instrument and the 
capabilities of the instrument itself introduce further modifications. 
In short, stellar spectra represent extremely complex, high-dimensional 
data that has gone through multiple transformations before reaching our 
eyes as a deceptively simple graph of flux versus wavelength.

Naturally, scientists want to extract as much information as possible from
a star's spectra to get back the properties that lie at the root of it all. 
However, this has proven to be no easy task. Spectroscopic surveys such as 
APOGEE develop their own in-house pipelines to extract stellar parameters
from spectra (ASPCAP in the case of APOGEE), and these pipelines are fairly
effective. However, they all work on the core assumption that the data to be 
analyzed is a list of pixels. In addition, the pipelines that surveys 
develop are closed-source and usually limited to the scope of the survey 
itself, causing a sort of fragmentation. This motivates the development of a 
more universal, open-source tool that can do what spectral analysis pipelines 
do, but in an alternative way.

Multiple efforts have been made in this direction, treating spectra in 
different ways. Some take the wavelength and flux values and construct
neural networks to create this mapping between stellar parameters and the
observed spectra, such as [EXAMPLE] and [EXAMPLE]. Others take a more 
physics-informed approach by decomposing spectra into some eigenbasis and 
treating it as a regression problem, such as \texttt{starfish}.
However, \texttt{blase} takes a different approach. It treats spectra not as 
a set of pixels or a set of eigenbasis coefficients but as a set of 
spectral lines, specifically Voigt profiles. There is no metric of this 
approach being better or worse than the others mentioned so far, but it is 
a different paradigm and is worth exploring as a semi-empirical tool for
spectral inference. This study aims to take advantage of \texttt{blase}'s
ability to change a spectrum's basis from flux versus wavelength into
spectral line versus spectral line properties, and use that to develop a
robust tool for stellar parameter inference.

However, this is only a proof of concept. Here we develop a tool that 
represents a very basic realization of the idea of line-by-line inference. 
We make no claims as to the performance of this tool, but rather aim to show 
that the idea is sound and able to be implemented, even if rudimentary, and 
that with further research and development, this tool has the potential to 
become useful for spectral inference.

We chose the PHOENIX synthetic spectral model grid as the basis for this 
study, because it is fairly well-known. While we plan to extend this tool to 
leverage information across multiple model grids in the future, both for 
accuracy and to increase the scope to include substellar objects with grids
such as Sonora, for now we limit our scope to a subset of the PHOENIX grid.

@@

\begin{table}[h]
    \hspace*{-1.1cm}\begin{tabular}{|c|c|c|}
        \hline
        \multicolumn{3}{|c|}{\textbf{PHOENIX Grid Subset}}\\
        \hline\hline
        Parameter & Symbol & Range\\
        \hline
        Alpha Element Abundance & $\alpha$ & [0] dex\\
        Iron Abundance & [Fe/H] & [-0.5, 0.5] dex\\
        Effective Temperature & $T_{\mathrm{eff}}$ & [2300, 12000] K\\
        Surface Gravity & $\log(g)$ & [0, 5]\\
        Wavelength & $\lambda$ & [8038, 12849] \AA\\
        \hline
    \end{tabular}
    \caption{The subset of the PHOENIX grid used in this study. These limits 
    were imposed to reduce the computational cost of the algorithms. The 
    wavelength limits in particular roughly line up with that of the 
    Habitable Zone Planet Finder (HPF) spectrograph.}
\end{table}

@@

Once optimization was complete, the list of identified lines, present in the
state dictionary of \texttt{blase}'s model, was saved to a \texttt{.pt} file
for each PHOENIX subset grid point.

@@

\subsection{Continuously Evaluable Manifolds}
For each line, the inputs to the interpolator were the three input parameters $T_{\mathrm{eff}}$,
$\log(g)$, and [Fe/H], and the output was a list of four parameters, 
$\delta\mu$, $\ln(a)$, $\sigma$, and $\gamma$. For each line, one of these 
interpolator objects was created using linear interpolation, and these 
interpolators were aggregated into a single list, which was when pickled 
and written to a \texttt{.pkl} file. These interpolators represent multiple 
manifolds mapping fundamental stellar parameters to spectral line properties. 
These interpolators could now be evaluated at any point lying within the 
domain of the PHOENIX subset, turning a discretely sampled PHOENIX subset 
into a continuous one. With the given size of the PHOENIX subset, the 
interpolator list takes up 19 GB of disk space. Since this is now evaluable,
we call this the PHOENIX generator.

\section{Bayesian Inference and Testing}
\subsection{Inference Algorithm}
With the PHOENIX generator complete, we used Bayesian optimization as the
inference algorithm, specifically the \texttt{gp\_minimize} function from
the \texttt{scikit-optimize} library. This algorithm uses a Gaussian Process
to model the objective function, which in this case was the $\chi^2$ loss
function between the interpolated spectrum and the true spectrum. The 
optimizer was configured to first run 10 random evaluations to seed the
surrogate model, then run 90 more evaluations now guided by the surrogate
model. This totals to 100 evaluations, which was deemed sufficient for this
study.

\subsection{Test Data}
Text

@@

\section{Discussion}
The computer used for this study, Triton, has the following specifications:
\begin{table}[h]
    \hspace*{0.7cm}\begin{tabular}{|c|c|}
        \hline
        \multicolumn{2}{|c|}{\textbf{Triton Specs}}\\
        \hline\hline
        CPU & AMD EPYC 7513\\
        Memory & 256 GB\\
        GPU & Nvidia A100 40GB ($\times$2)\\
        \hline
    \end{tabular}
    \caption{This computer is owned by Caroline Morley and is available to
    members of her research group. It was used for all computations, but not 
    for generating visualizations.}
\end{table}

@@

    \item \textit{Low-Order Interpolation}: The current design of the 
    algorithm uses linear interpolation to construct manifolds, however this
    could be made more sophisticated with more complex interpolation schemes.
    However, the drawback of this is that higher orders of interpolation
    become increasingly expensive both in terms of disk utilization, and 
    in terms of the computational cost to generate and evaluate them. Future
    studies should sidestep this issue by using more generalizable methods
    that are much more efficient.
